{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2017-10-20 sftrends_manu.Rnw
some tables/figs back to supplement
no footnotes in tables, changed color sheme of clam plot

@@

The results for Suisun Bay provide additional descriptions of change in production as it relates to flow, grazing, and nitrogen ratios.  In general, clam biomass was associated with a decrease in \ac{chl} concentration, as shown by others \citep{Alpin92,Thompson08}. Our results also suggested that diatoms were the dominant genera early in the time series, particularly in late summer, whereas the spring peak observed in later years represents a shift to an earlier seasonal maxima.  This supports past research that showed a decrease in silica uptake by diatoms following invasion \citep{Cloern96,Kimmerer05}.  Anontrivial portion of the \ac{din} increase could be related to the decrease in a major `sink', i.e., decreased \ac{din} uptake by phytoplankton due to top down grazing pressure from \textit{P. amurensis}.  Flow effects on phytoplankton production have also changed over time. In the absence of benthic grazing prior to invasion, \ac{chla} production was limited at low flow as less nutrients were exported from the Delta, stimulated as flow increases, and reduced at high flow as either nutrients or phytoplankton biomass are exported to the estuary (\cref{fig:clmchl}j). Recent years have shown a decrease in overall chlorophyll, with particularly low concentrations at low flow (high salinity).  As such, \ac{chla} production in early years is directly related to flow, whereas the relationship with flow in later years is indirect as increased flow reduces clam abundance and releases phytoplankton from benthic grazing pressure. These relationships have been suggested by others \citep{Cloern83,Alpine92,Parchaso02,Jassby08}, although the precise mechanism demonstrated by \ac{wrtds} provides a quantitative description of factors that drive water quality in the Delta.

@@

##
# annual clam time series
p1 <- ggplot(clmdat, aes(x = yr, y = clams_smp, fill = species)) + 
  geom_bar(stat = 'identity') + 
  mythm +
  scale_fill_brewer(
    palette = 'Set2',
    labels = parse(text = c('italic(Corbicula)', 'italic(Potamocorbula)'))
    ) + 
  scale_x_date('Year', limits = xlims) +
  scale_y_continuous('Clams per sample') +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  labs(title = '(a)')

@@

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and annual aggregations'
foot.val<-'\\footnotesize Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. *$p<0.05$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Annual'),
  n.cgroup = c(2),
  rowname = stats,
  label = 'tab:trndsann',
  insert.bottom = foot.val
  )

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1976-1995'
foot.val<-'\\footnotesize Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Seasonal, 1976-1995'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsbef',
  insert.bottom = foot.val
  )

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1996-2013'
foot.val<-'\\footnotesize Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Seasonal, 1996-2013'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsaft',
  insert.bottom = foot.val
  )

@@

cap.val<-'Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \\ac{wwtp} upgrades upstream of station P8'
foot.val<-'\\footnotesize Upgrades were completed in 2006 at the City of Stockton \\ac{wwtp} (San Joaquin County, \\cref{fig:stock}). Summaries are  medians and percent change per year in parentheses (increasing in bold).  Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(2, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2),
  rowname = cats,
  label = 'tab:p8chg',
  insert.bottom = foot.val
  )

@@

>>>>

%for supplemental figures/tables
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

@@

The results for Suisun Bay provide additional descriptions of change in production as it relates to flow, grazing, and nitrogen ratios.  In general, clam biomass was associated with a decrease in \ac{chla} concentration, as shown by others \citep{Alpine92,Thompson08}. Our results also suggested that diatoms were the dominant genera early in the time series, particularly in late summer, whereas the spring peak observed in later years represents a shift to an earlier seasonal maxima.  This supports past research that showed a decrease in silica uptake by diatoms following invasion \citep{Cloern96,Kimmerer05}.  Anontrivial portion of the \ac{din} increase could be related to the decrease in a major `sink', i.e., decreased \ac{din} uptake by phytoplankton due to top down grazing pressure from \textit{P. amurensis}.  Flow effects on phytoplankton production have also changed over time. In the absence of benthic grazing prior to invasion, \ac{chla} production was limited at low flow as less nutrients were exported from the Delta, stimulated as flow increases, and reduced at high flow as either nutrients or phytoplankton biomass are exported to the estuary (\cref{fig:clmchl}j). Recent years have shown a decrease in overall chlorophyll, with particularly low concentrations at low flow (high salinity).  As such, \ac{chla} production in early years is directly related to flow, whereas the relationship with flow in later years is indirect as increased flow reduces clam abundance and releases phytoplankton from benthic grazing pressure. These relationships have been suggested by others \citep{Cloern83,Alpine92,Parchaso02,Jassby08}, although the precise mechanism demonstrated by \ac{wrtds} provides a quantitative description of factors that drive water quality in the Delta.

@@

##
# annual clam time series
p1 <- ggplot(clmdat, aes(x = yr, y = clams_smp, fill = species)) + 
  geom_bar(stat = 'identity', colour = 'black') + 
  mythm +
  scale_fill_manual(
    values = c('white', 'lightgrey'),
    labels = parse(text = c('italic(Corbicula)', 'italic(Potamocorbula)'))
    ) + 
  scale_x_date('Year', limits = xlims) +
  scale_y_continuous('Clams per sample') +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  labs(title = '(a)')

@@

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and annual aggregations. Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. *$p<0.05$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Annual'),
  n.cgroup = c(2),
  rowname = stats,
  label = 'tab:trndsann'
  )

@@

cap.val<-'Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \\ac{wwtp} upgrades upstream of station P8. Upgrades were completed in 2006 at the City of Stockton \\ac{wwtp} (San Joaquin County, \\cref{fig:stock}). Summaries are  medians and percent change per year in parentheses (increasing in bold).  Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(2, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2),
  rowname = cats,
  label = 'tab:p8chg'
  )

@@

\beginsupplement

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1976-1995. Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Seasonal, 1976-1995'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsbef'
  )

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1996-2013. Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Seasonal, 1996-2013'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsaft'
  )

### 2017-10-20 sftrends_manu.tex
some tables/figs back to supplement
no footnotes in tables, changed color sheme of clam plot

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^{1*}$, Thomas W. Jabusch$^2$, Philip R. Trowbridge$^2$, David B. Senn$^2$}}
  \\\\{\textit {\normalsize $^*$Corresponding author: \href{mailto:marcusb@sccwrp.org}{marcusb@sccwrp.org}}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
  \\\\{\textit {\normalsize Current address: Southern California Coastal Water Research Project}}
  \\{\textit {\normalsize 3535 Harbor Blvd, Suite 110, Costa Mesa, CA 92626}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Ave, Richmond, CA 94804}}
  \vspace{1in}
  \\ Version Date:   Thu Oct 19 15:58:58 2017 -0700
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

% overall trends, years
%latex.default(tab[, -c(1, 2)], file = "", rowlabel = "Analyte/Station",     caption = cap.val, caption.loc = "top", rgroup = vars, n.rgroup = rep(nrow(tab)/3,         3), cgroup = c("Annual"), n.cgroup = c(2), rowname = stats,     label = "tab:trndsann", insert.bottom = foot.val)%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrogen analytes for all stations and annual aggregations\label{tab:trndsann}} 
\begin{center}
\begin{tabular}{lll}
\hline\hline
\multicolumn{1}{l}{\bfseries Analyte/Station}&\multicolumn{2}{c}{\bfseries Annual}\tabularnewline
\cline{2-3}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{1976-1995}&\multicolumn{1}{c}{1996-2013}\tabularnewline
\hline
{\bfseries DIN}&&\tabularnewline
~~C10&1.3 \footnotesize{(\textbf{0.8})*}&1.4 \footnotesize{(-3.1)*}\tabularnewline
~~C3&0.3 \footnotesize{(\textbf{2.2})*}&0.5 \footnotesize{(-0.1)*}\tabularnewline
~~D19&0.4 \footnotesize{(\textbf{0.2})*}&0.4 \footnotesize{(-1.9)*}\tabularnewline
~~D26&0.4 \footnotesize{(\textbf{0.4})*}&0.5 \footnotesize{(-1.2)*}\tabularnewline
~~D28&0.4 \footnotesize{(\textbf{0.1})*}&0.4 \footnotesize{(-3.1)*}\tabularnewline
~~D4&0.3 \footnotesize{(\textbf{0.6})*}&0.4 \footnotesize{(-0.3)*}\tabularnewline
~~D6&0.4 \footnotesize{(\textbf{1.8})*}&0.5 \footnotesize{(-0.3)*}\tabularnewline
~~D7&0.4 \footnotesize{(\textbf{1.7})*}&0.5 \footnotesize{(-0.7)*}\tabularnewline
~~MD10&0.4 \footnotesize{(-1.1)*}&0.3 \footnotesize{(-2.4)*}\tabularnewline
~~P8&1.3 \footnotesize{(\textbf{2.5})*}&1.7 \footnotesize{(-2)*}\tabularnewline
\hline
{\bfseries NH$_{4}^{+}$}&&\tabularnewline
~~C10&0.1 \footnotesize{(-3.4)*}&0 \footnotesize{(-5.2)*}\tabularnewline
~~C3&0.2 \footnotesize{(\textbf{3.7})*}&0.3 \footnotesize{(\textbf{0})}\tabularnewline
~~D19&0 \footnotesize{(\textbf{0.4})*}&0 \footnotesize{(-1.7)*}\tabularnewline
~~D26&0.1 \footnotesize{(\textbf{2.2})*}&0.1 \footnotesize{(-1.5)*}\tabularnewline
~~D28&0 \footnotesize{(-1.1)*}&0 \footnotesize{(-1.4)*}\tabularnewline
~~D4&0 \footnotesize{(\textbf{0.9})*}&0.1 \footnotesize{(\textbf{0})}\tabularnewline
~~D6&0.1 \footnotesize{(\textbf{2.4})*}&0.1 \footnotesize{(-0.5)*}\tabularnewline
~~D7&0.1 \footnotesize{(\textbf{1.5})*}&0.1 \footnotesize{(-1.2)*}\tabularnewline
~~MD10&0.1 \footnotesize{(-2.8)*}&0 \footnotesize{(-1.1)*}\tabularnewline
~~P8&0.2 \footnotesize{(\textbf{4.9})*}&0.1 \footnotesize{(-10.3)*}\tabularnewline
\hline
{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&&\tabularnewline
~~C10&1.2 \footnotesize{(\textbf{1.4})*}&1.4 \footnotesize{(-3)*}\tabularnewline
~~C3&0.1 \footnotesize{(-0.1)*}&0.2 \footnotesize{(\textbf{0.7})*}\tabularnewline
~~D19&0.4 \footnotesize{(-0.1)*}&0.4 \footnotesize{(-1.9)*}\tabularnewline
~~D26&0.3 \footnotesize{(\textbf{0})}&0.4 \footnotesize{(-1.1)*}\tabularnewline
~~D28&0.4 \footnotesize{(-0.2)*}&0.4 \footnotesize{(-3.1)*}\tabularnewline
~~D4&0.3 \footnotesize{(\textbf{0.7})*}&0.3 \footnotesize{(-0.4)*}\tabularnewline
~~D6&0.3 \footnotesize{(\textbf{1.3})*}&0.4 \footnotesize{(-0.3)*}\tabularnewline
~~D7&0.4 \footnotesize{(\textbf{0.7})*}&0.4 \footnotesize{(-0.7)*}\tabularnewline
~~MD10&0.4 \footnotesize{(-1)*}&0.3 \footnotesize{(-2.5)*}\tabularnewline
~~P8&1.2 \footnotesize{(\textbf{1.7})*}&1.5 \footnotesize{(-0.6)*}\tabularnewline
\hline
\end{tabular}\end{center}
\footnotesize Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. *$p<0.05$\end{table}

@@

%latex.default(tab[, -c(1, 2)], file = "", rowlabel = "Analyte/Station",     caption = cap.val, caption.loc = "top", rgroup = vars, n.rgroup = rep(nrow(tab)/3,         3), cgroup = c("Seasonal, 1976-1995"), n.cgroup = c(4),     rowname = stats, label = "tab:trndsbef", insert.bottom = foot.val)%

@@

\caption{Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1976-1995\label{tab:trndsbef}} 

@@

\footnotesize Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$\end{table}

@@

%latex.default(tab[, -c(1, 2)], file = "", rowlabel = "Analyte/Station",     caption = cap.val, caption.loc = "top", rgroup = vars, n.rgroup = rep(nrow(tab)/3,         3), cgroup = c("Seasonal, 1996-2013"), n.cgroup = c(4),     rowname = stats, label = "tab:trndsaft", insert.bottom = foot.val)%

@@

\caption{Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1996-2013\label{tab:trndsaft}} 

@@

\footnotesize Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$\end{table}

% trends in no23, nh at p8
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(2, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2), rowname = cats,     label = "tab:p8chg", insert.bottom = foot.val)%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \ac{wwtp} upgrades upstream of station P8\label{tab:p8chg}} 
\begin{center}
\begin{tabular}{lllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NH$_{4}^{+}$}\tabularnewline
\cline{2-3} \cline{5-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries Annual}&&&&&\tabularnewline
~~1976-2006&1.3&\textbf{2}*&&0.2&\textbf{2.8}*\tabularnewline
~~2007-2013&1.4&-1.9*&&0.1&-16.6*\tabularnewline
\hline
{\bfseries Seasonal, pre}&&&&&\tabularnewline
~~Spring&1.2&\textbf{1.6}*&&0.2&\textbf{1.4}*\tabularnewline
~~Summer&1&\textbf{2.4}*&&0.1&\textbf{3.3}*\tabularnewline
~~Fall&1.3&\textbf{2.2}*&&0.2&\textbf{4.9}*\tabularnewline
~~Winter&1.5&\textbf{2.1}*&&0.7&\textbf{4.8}*\tabularnewline
\hline
{\bfseries Seasonal, post}&&&&&\tabularnewline
~~Spring&1.3&-1.6*&&0.1&-16.2*\tabularnewline
~~Summer&0.9&-4.3*&&0.1&-15.7*\tabularnewline
~~Fall&1.5&-1.7*&&0.1&-19.3*\tabularnewline
~~Winter&2.2&-0.8*&&0.2&-26.7*\tabularnewline
\hline
\end{tabular}\end{center}
\footnotesize Upgrades were completed in 2006 at the City of Stockton \ac{wwtp} (San Joaquin County, \cref{fig:stock}). Summaries are  medians and percent change per year in parentheses (increasing in bold).  Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$\end{table}

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^{1*}$, Thomas W. Jabusch$^2$, Philip R. Trowbridge$^2$, David B. Senn$^2$}}
  \\\\{\textit {\normalsize $^*$Corresponding author: \href{mailto:marcusb@sccwrp.org}{marcusb@sccwrp.org}}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
  \\\\{\textit {\normalsize Current address: Southern California Coastal Water Research Project}}
  \\{\textit {\normalsize 3535 Harbor Blvd, Suite 110, Costa Mesa, CA 92626}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Ave, Richmond, CA 94804}}
  \vspace{1in}
  \\ Version Date:   Fri Oct 20 11:14:05 2017 -0700
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

% overall trends, years
%latex.default(tab[, -c(1, 2)], file = "", rowlabel = "Analyte/Station",     caption = cap.val, caption.loc = "top", rgroup = vars, n.rgroup = rep(nrow(tab)/3,         3), cgroup = c("Annual"), n.cgroup = c(2), rowname = stats,     label = "tab:trndsann")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrogen analytes for all stations and annual aggregations. Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. *$p<0.05$\label{tab:trndsann}} 
\begin{center}
\begin{tabular}{lll}
\hline\hline
\multicolumn{1}{l}{\bfseries Analyte/Station}&\multicolumn{2}{c}{\bfseries Annual}\tabularnewline
\cline{2-3}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{1976-1995}&\multicolumn{1}{c}{1996-2013}\tabularnewline
\hline
{\bfseries DIN}&&\tabularnewline
~~C10&1.3 \footnotesize{(\textbf{0.8})*}&1.4 \footnotesize{(-3.1)*}\tabularnewline
~~C3&0.3 \footnotesize{(\textbf{2.2})*}&0.5 \footnotesize{(-0.1)*}\tabularnewline
~~D19&0.4 \footnotesize{(\textbf{0.2})*}&0.4 \footnotesize{(-1.9)*}\tabularnewline
~~D26&0.4 \footnotesize{(\textbf{0.4})*}&0.5 \footnotesize{(-1.2)*}\tabularnewline
~~D28&0.4 \footnotesize{(\textbf{0.1})*}&0.4 \footnotesize{(-3.1)*}\tabularnewline
~~D4&0.3 \footnotesize{(\textbf{0.6})*}&0.4 \footnotesize{(-0.3)*}\tabularnewline
~~D6&0.4 \footnotesize{(\textbf{1.8})*}&0.5 \footnotesize{(-0.3)*}\tabularnewline
~~D7&0.4 \footnotesize{(\textbf{1.7})*}&0.5 \footnotesize{(-0.7)*}\tabularnewline
~~MD10&0.4 \footnotesize{(-1.1)*}&0.3 \footnotesize{(-2.4)*}\tabularnewline
~~P8&1.3 \footnotesize{(\textbf{2.5})*}&1.7 \footnotesize{(-2)*}\tabularnewline
\hline
{\bfseries NH$_{4}^{+}$}&&\tabularnewline
~~C10&0.1 \footnotesize{(-3.4)*}&0 \footnotesize{(-5.2)*}\tabularnewline
~~C3&0.2 \footnotesize{(\textbf{3.7})*}&0.3 \footnotesize{(\textbf{0})}\tabularnewline
~~D19&0 \footnotesize{(\textbf{0.4})*}&0 \footnotesize{(-1.7)*}\tabularnewline
~~D26&0.1 \footnotesize{(\textbf{2.2})*}&0.1 \footnotesize{(-1.5)*}\tabularnewline
~~D28&0 \footnotesize{(-1.1)*}&0 \footnotesize{(-1.4)*}\tabularnewline
~~D4&0 \footnotesize{(\textbf{0.9})*}&0.1 \footnotesize{(\textbf{0})}\tabularnewline
~~D6&0.1 \footnotesize{(\textbf{2.4})*}&0.1 \footnotesize{(-0.5)*}\tabularnewline
~~D7&0.1 \footnotesize{(\textbf{1.5})*}&0.1 \footnotesize{(-1.2)*}\tabularnewline
~~MD10&0.1 \footnotesize{(-2.8)*}&0 \footnotesize{(-1.1)*}\tabularnewline
~~P8&0.2 \footnotesize{(\textbf{4.9})*}&0.1 \footnotesize{(-10.3)*}\tabularnewline
\hline
{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&&\tabularnewline
~~C10&1.2 \footnotesize{(\textbf{1.4})*}&1.4 \footnotesize{(-3)*}\tabularnewline
~~C3&0.1 \footnotesize{(-0.1)*}&0.2 \footnotesize{(\textbf{0.7})*}\tabularnewline
~~D19&0.4 \footnotesize{(-0.1)*}&0.4 \footnotesize{(-1.9)*}\tabularnewline
~~D26&0.3 \footnotesize{(\textbf{0})}&0.4 \footnotesize{(-1.1)*}\tabularnewline
~~D28&0.4 \footnotesize{(-0.2)*}&0.4 \footnotesize{(-3.1)*}\tabularnewline
~~D4&0.3 \footnotesize{(\textbf{0.7})*}&0.3 \footnotesize{(-0.4)*}\tabularnewline
~~D6&0.3 \footnotesize{(\textbf{1.3})*}&0.4 \footnotesize{(-0.3)*}\tabularnewline
~~D7&0.4 \footnotesize{(\textbf{0.7})*}&0.4 \footnotesize{(-0.7)*}\tabularnewline
~~MD10&0.4 \footnotesize{(-1)*}&0.3 \footnotesize{(-2.5)*}\tabularnewline
~~P8&1.2 \footnotesize{(\textbf{1.7})*}&1.5 \footnotesize{(-0.6)*}\tabularnewline
\hline
\end{tabular}\end{center}
\end{table}

% trends in no23, nh at p8
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(2, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2), rowname = cats,     label = "tab:p8chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \ac{wwtp} upgrades upstream of station P8. Upgrades were completed in 2006 at the City of Stockton \ac{wwtp} (San Joaquin County, \cref{fig:stock}). Summaries are  medians and percent change per year in parentheses (increasing in bold).  Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$\label{tab:p8chg}} 
\begin{center}
\begin{tabular}{lllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NH$_{4}^{+}$}\tabularnewline
\cline{2-3} \cline{5-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries Annual}&&&&&\tabularnewline
~~1976-2006&1.3&\textbf{2}*&&0.2&\textbf{2.8}*\tabularnewline
~~2007-2013&1.4&-1.9*&&0.1&-16.6*\tabularnewline
\hline
{\bfseries Seasonal, pre}&&&&&\tabularnewline
~~Spring&1.2&\textbf{1.6}*&&0.2&\textbf{1.4}*\tabularnewline
~~Summer&1&\textbf{2.4}*&&0.1&\textbf{3.3}*\tabularnewline
~~Fall&1.3&\textbf{2.2}*&&0.2&\textbf{4.9}*\tabularnewline
~~Winter&1.5&\textbf{2.1}*&&0.7&\textbf{4.8}*\tabularnewline
\hline
{\bfseries Seasonal, post}&&&&&\tabularnewline
~~Spring&1.3&-1.6*&&0.1&-16.2*\tabularnewline
~~Summer&0.9&-4.3*&&0.1&-15.7*\tabularnewline
~~Fall&1.5&-1.7*&&0.1&-19.3*\tabularnewline
~~Winter&2.2&-0.8*&&0.2&-26.7*\tabularnewline
\hline
\end{tabular}\end{center}
\end{table}

\beginsupplement

@@

%latex.default(tab[, -c(1, 2)], file = "", rowlabel = "Analyte/Station",     caption = cap.val, caption.loc = "top", rgroup = vars, n.rgroup = rep(nrow(tab)/3,         3), cgroup = c("Seasonal, 1976-1995"), n.cgroup = c(4),     rowname = stats, label = "tab:trndsbef")%

@@

\caption{Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1976-1995. Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$\label{tab:trndsbef}} 

@@

\end{table}

@@

%latex.default(tab[, -c(1, 2)], file = "", rowlabel = "Analyte/Station",     caption = cap.val, caption.loc = "top", rgroup = vars, n.rgroup = rep(nrow(tab)/3,         3), cgroup = c("Seasonal, 1996-2013"), n.cgroup = c(4),     rowname = stats, label = "tab:trndsaft")%

@@

\caption{Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1996-2013. Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF. *$p<0.05$\label{tab:trndsaft}} 

@@

\end{table}
